Vector Representations of Linguistic Communities and Associated Models

### 2021-12-15 paper/paper1.tex
a
revise up to procrustes

\title{Community-conditioned language models for linguistic variation}

@@

\begin{document}
\maketitle
\begin{abstract}
  Community-level linguistic variation is a core concept in sociolinguistics,
  but more tools are needed to study variation from a computational perspective.
  In this paper, we use conditional language models to investigate linguistic variation
  across 510 online communities.
  We probe the community representations learned by the conditional language model
  and find that they are highly correlated with a social network-based 
  representation that does not use any linguistic input.
\end{abstract}

@@

Linguistic communication requires that speakers share
certain linguistic conventions, such as syntactic
structure, word meanings, and patterns of interaction.
Speakers assume that these conventions are \emph{common ground} among
their interlocutors, based on joint membership in a community
\cite{Stalnaker2002, Clark1996}.  Such \emph{speech communities} \citep{Gumperz1972} 
range in size from the very small, like members of a friend group, 
to the very large, like speakers of English; however as \citet{Eckert1992}
point out, it is \emph{communities of practice}---defined by mutual social engagement
in a common activity---that are the primary locus of linguistic variation.

Variation is an important object of study in sociolinguistics,
and is naturally suited for computational analysis \citep{Nguyen2016}.
Most previous computational work on linguistic variation 
has considered variation at the level of macro-social categories, such as gender
\citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
geographic location \citep{Eisenstein2010,Bamman2014a}.%
\footnote{Recent exceptions are \citet{DelTredici2017} who use a modified
skip-gram model to show that lexical semantic variation occurs even across 
different communities organised around the same topic and 
\citet{Lucy2021}, who use contextualised word representations to
investigate community-level variation in word sense distributions.}
In the present work, we investigate linguistic variation
in online communities on the social media website Reddit. 
In contrast to previous work, we investigate variation across 
speech communities, rather than macro-social categories.

@@

For this purpose, we introduce (\cref{sec:cclm}) various
Community-Conditioned Language Models (CCLMs for short). These models
are conditioned on a vector representation (or embedding), which
varies for each community. Hence, they learn \emph{community
  embeddings}.  Although we test which architectures best make use
of the community information, our purpose is not to improve models in
terms of perplexity, but rather we want to extract the best set of
community embeddings and test how the resulting embeddings
predict the social structure of subreddits. For this purpose, we test
the correlation with a social network-based representation of
communities (\cref{sec:embedding-analysis}).

@@

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}. Although Transformer-based language models are considered
state of the art, they achieve dominance partly thanks to the
availability of very large data sets \citep[e.g.,][]{Devlin2019,Brown2020},
which are not available to us.\footnote{Fine-tuning existing models
  is not compatible with our methodology, since we would fundamentally
  change the structure of the network by concatenating
  community embeddings.}

@@

Messages were preprocessed as follows:
We excluded the content of block quotes, code blocks, and tables and
removed markup (formatting) commands, 
extracting only rendered text.
Messages were tokenized using the default English model for the SpaCy tokenizer 
Version 2.2.3 \citep{Honnibal2017}.

@@

We trained the models until the validation loss stopped decreasing for
two epochs in a row, and used the weights from the epoch with the
smallest validation loss for testing.

@@

We observe empirically that across all the models we
tested, communities where conditioning has the least effect tend to be
organised around more general interest topics, such as 
\subreddit{relationships} and \subreddit{advice}, where the subject
matter is relevant to a broad range of people.  Conditioning the model
on community appears to have the most benefit for narrower special-interest
subreddits, such as those organised around a certain videogame, sports
team, or subculture.  This supports the idea that communities of practice
are the primary locus of linguistic variation.
%makes sense intuitively, since communities
%with more niche topics would tend to have more specialised vocabulary
%and other linguistic patterns.

@@

A difficulty is that the social embeddings and linguistic embeddings
range over completely unrelated spaces, so one cannot use the usual
cosine similarity metric \emph{across} this boundary. One can,
however, use cosine similarity between \emph{pairs} of communities,
and verify the similarities correlated between linguistic to social
embeddings.  This gives us a way of characterizing the differences
between the two kinds of community representation. To get a more
concrete sense of what this method yields, we first survey some of the
most salient community pairs. We stress that this survey is not meant
as a rigorous statistical analysis, as we shall see. Rather it is
meant to give a flavor of discrepancies and similarities existing
between linguistic and social relations.

@@

We simultaneously have examples of high correlation between social and
linguistic embeddings, but also of low correlation. This is
essentially anecdotal evidence: we must resort to statistical tools to
extract the general trends.
A straightforward, but ultimately flawed,
way to measure how similar the two spaces are is to consider each pair
of communities \((i,j)\), and compute the correlation between the
cosine similarities of both embeddings.

>>>>

\title{Vector Representations of Linguistic Communities and Associated Models}
 % Representations -conditioned language models for linguistic variation

@@

\begin{document}
\maketitle
\begin{abstract}
  Community-level linguistic variation is a core concept in sociolinguistics.
  % but more tools are needed to study variation from a computational perspective.
  In this paper, we use computational methods to investigate how much linguistic variation corresponds with social connections between communities.
  Technically, use conditional language models to investigate linguistic variation
  across 510 online communities.
  We probe the community representations learned by the conditional language model
  and find that they are highly correlated with a social network-based 
  representation that does not use any linguistic input.
\end{abstract}

@@

Linguistic communication requires that speakers share
certain linguistic conventions, such as syntactic
structure, word meanings, and patterns of interaction.
Speakers assume that these conventions are \emph{common ground} among
their interlocutors, based on joint membership in a community
\cite{Stalnaker2002, Clark1996}.  Such \emph{speech communities} \citep{Gumperz1972} 
range in size from the very small, like members of a friend group, 
to the very large, like speakers of English. However, as \citet{Eckert1992}
point out, it is \emph{communities of practice}---defined by mutual social engagement
in a common activity---that are the primary locus of linguistic variation.

Variation is an important object of study in sociolinguistics,
and is naturally amenable to computational analysis \citep{Nguyen2016}.
Most previous computational work on linguistic variation 
has considered variation at the level of macro-social categories, such as gender
\citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
geographic location \citep{Eisenstein2010,Bamman2014a}.%
\footnote{Recent exceptions are \citet{DelTredici2017} who use a modified
skip-gram model to show that lexical semantic variation occurs even across 
different communities organised around the same topic and 
\citet{Lucy2021}, who use contextualised word representations to
investigate community-level variation in word sense distributions.}
In the present work, we investigate linguistic variation
in online communities on the social media website Reddit. 
This contrasts with previous work: we investigate variation across 
speech communities, rather than macro-social categories.

@@

For this purpose, we introduce (\cref{sec:cclm}) various
Community-Conditioned Language Models (CCLMs for short). These models
are conditioned on a vector representation (or embedding), which
varies for each community. Hence, they learn \emph{community
  embeddings}.  Even though we report which architectures make best use
of the community information, our primary purpose is not to improve models in
terms of perplexity, but rather we want to extract the best set of
community embeddings and test how the resulting embeddings
predict the social structure of subreddits. Technically, we test
the correlation with a social network-based representation of
communities (\cref{sec:embedding-analysis}).

@@

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}. Although Transformer-based language models are considered
state-of-the-art, they achieve dominance partly thanks to the
availability of very large data sets \citep[e.g.,][]{Devlin2019,Brown2020},
which are not available to us.\footnote{Fine-tuning existing models
  is not compatible with our methodology, because we fundamentally
  change the structure of the network by concatenating
  community embeddings with hidden states at various levels.} Thus the LSTM is a worthy model to test for us.

@@

Messages were preprocessed as follows:
we excluded the content of block quotes, code blocks, and tables and
removed markup (formatting) commands, 
extracting only rendered text.
Messages were tokenized using the default English model for the SpaCy tokenizer 
Version 2.2.3 \citep{Honnibal2017}.

@@

We trained the models until the validation loss stopped decreasing for
two epochs in a row, and used the weights from the epoch with the
smallest validation loss for testing. 
Each training epoch took approximately 1.5 hours of GPU time.

@@

We observe that across all the models we
tested, communities where conditioning has the least effect tend to be
organised around more general interest topics, such as 
\subreddit{relationships} and \subreddit{advice}, where the subject
matter is relevant to a broad range of people.  Conditioning the model
on community appears to have the most benefit for narrower special-interest
subreddits, such as those organised around a certain videogame, sports
team, or subculture.  These empirical observations corroborate the idea that communities of practice
are the primary locus of linguistic variation.
%makes sense intuitively, since communities
%with more niche topics would tend to have more specialised vocabulary
%and other linguistic patterns.

@@

When comparing social embeddings and linguistic embeddings, a difficulty is that they
range over completely unrelated spaces. Thus one cannot use the usual
cosine similarity metric \emph{between} these spaces. One can,
however, use cosine similarity between \emph{pairs} of communities,
and verify that the similarities are correlated between linguistic and social
embeddings.  This gives a way of characterizing the differences
between the two kinds of community representation. To get a more
concrete sense of what this method yields, we first survey some of the
most salient community pairs. We stress that this survey is not meant
as a rigorous statistical analysis, as we shall see. Rather it is
meant to give a flavor of discrepancies and similarities existing
between linguistic and social relations.

@@

In sum, empirical observation simultaneously reveals examples of high
and low correlation between social and linguistic embeddings.  To
quantify correlation and extract the general trends, we must resort to
statistical tools, as we do below.

A straightforward (but ultimately flawed) way to measure how similar
the two spaces are would be to generalise the above method, by
consider each pair of communities \((i,j)\), and compute the
correlation between the cosine similarities of both embeddings.
